#Excercise Activity Analysis
By Frank Pinto
---

##Reading in and cleaning the Data
```{r}
dat <- read.csv("activity.csv")
cdat <- dat[complete.cases(dat),]
library(plyr)
library(dplyr)

```

##Histogram
```{r}
sumsteps <- ddply(cdat, "date", .fun = function(x) sum(x$steps))
hist(sumsteps[,2], xlab = "Date", ylab = "Steps Taken", main = "Steps taken by Date")
```

## Mean and Median of Steps taken each day

The Median is
```{r}
median(sumsteps[,2])
```
The Mean is
```{r}
mean(sumsteps[,2])
```

##Average pattern per day
```{r}
avgsteps <- ddply(cdat, "interval", .fun = function(x) mean(x$steps))
plot(x = avgsteps[,1], y = avgsteps[,2], xlab = "Interval", ylab = "Mean", main = "Mean of Steps taken by Interval")
```

The Interval with the most steps taken is
```{r}
avgsteps[which.max(avgsteps[,2]),]
```

##NAs found in the Data
The number of NAs found in the Data is
```{r}
sum(!complete.cases(dat))
```

Missing data will be filled with the mean for their corresponding interval
```{r}
a <- dat[is.na(dat),]
b <- ddply(cdat, "interval", function(x) mean(x$steps))
fread <- replace(a$steps, a$interval %in% b$interval, b$V1)
a[,1] <- fread 
dat[is.na(dat),] <- a
str(dat)
```

Calculating the amount of NAs in the new DT
```{r}
sum(!complete.cases(dat))
```

The Mean of the Steps taken by Date of the new dat DT
```{r}
sumsteps <- ddply(dat, "date", .fun = function(x) sum(x$steps))
mean(sumsteps$V1)
```
The Median of the Steps taken by Date of the new dat DT
```{r}
median(sumsteps$V1)
```
The Histogram of the Steps taken by Date of the new dat DT
```{r}
hist(sumsteps[,2], xlab = "Date", ylab = "Steps Taken", main = "Steps taken by Date")
```

Given the changes in the dat DT the mean went up by 1 step, while the median remained the same. The Histogram showed notorious change in the highest bar of aprox 10.

##Weekdays vs. Weekends

Setting the Date column as factors
```{r}
dat$date <- weekdays(as.Date(dat$date))
dat$date <- as.factor(dat$date)
str(dat$date)
dat$date <- revalue(dat$date, c("domingo"="Weekend", "sábado"="Weekend"))
dat$date <- revalue(dat$date, c("lunes"="Weekday", "martes"="Weekday", "miércoles"="Weekday", "jueves"="Weekday", "viernes"="Weekday"))
str(dat)
```

Plot of steps taken in each interval during Weekends (bottom) and Weekdays (top)
```{r}
a <- a <- dat[dat$date == "Weekday",]
a <- ddply(a, "interval", function(x) mean(x$steps))
b <- dat[dat$date == "Weekend",]
b <- ddply(b, "interval", function(x) mean(x$steps))
par(mfrow = c(2,1))
par(mar = c(0,0,0,0), oma = c(4,4,0,0))
plot(x = a$interval, y = a$V1)
plot(x = b$interval, y = b$V1)

